DeepSphere: a graph-based spherical CNN

### 2019-09-25 main.tex
cosmo cost-accuracy figure

\usepackage{hyperref}
\usepackage{url}
\usepackage{breqn}
\usepackage{booktabs}
\usepackage{siunitx} % used in SHREC table
\usepackage{bbm}        % for mathbbm
\usepackage{graphicx} \graphicspath{{figures/}}
\usepackage[export]{adjustbox} % figure vertical alignment
\usepackage{subfig}
%\usepackage{wrapfig}
%\usepackage{floatflt}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{wrapfig}

@@

\todo{
\begin{itemize}
	\item shall we really name it? Or simply reference the general idea of a graph-based spherical CNN.
	\item adjacency matrix: $A$ instead of $W$?
	\item anonymize DeepSphere citation
% 	\item climate spherical figure: inputs with TC and AR labels superimposed [Frédérick]
	\item citet for inline, citep for parenthesis
	\item More details on related work? Not sure we have the space. If yes, detailed comparison with related work after method.
	\item PSD of climate data: how does it compare to cosmo and 3D objects?
	\item align the spheres in figure 1
	\item mention that \citep{esteves2018sphericalcnn} and \citep{cohen2018sphericalcnn} computed equivariance error
	\item check footnotes
	\item climate data on spectrum plot
	\item result tables: only what supports the main message in the text, everything we have in the appendix
	\item Unify vocabulary. Sampling, discretization, or pixelization of the sphere? Signal, maps, functions. Irregular, uneven. Coarsening, down-sampling.
	\item check English vs American English
	\item unify enumerations. I like (i), (ii), etc.
	\item notation: should we use indices from 1 to n instead of 0 to n-1 for compactness?
	\item add \citet{bruna2013gnn} in equivariance? They were the first to do a spherical CNN with graphs. Graph is built from a uniform random sampling of the sphere.
	\item figure placements
	\item do the experiments on the icosahedral grid with the pooling as defined in the paper
	\item put all the code (learning experiments from Frédérick and equivariance experiments from Martino) in a common reproducible git repository to be linked from the paper
	\item update pygsp and merge sphere branch
	\item check that NN architectures and hyper-params are correctly reported
	\item keep NN architecture and hyper-params in the main text? Move the paragraph to appendix if not enough space.
	%\item shall we still site adam and batch norm? no, mainstream enough
	\item check results for ModelNet40 and integrate in text
	\item Check that HEALPix is introduced in the introduction
	\item heuristic used to set the kernel width in v1. Add values to \figref{kernel_widths} for comparison.
	\item check experiments section [Frédérick]: all the results numbers and NN + training details
\end{itemize}
}
\newpage

@@

Given observations, cosmologists estimate the posterior probability of cosmological parameters, such as the matter density $\Omega_m$ and the normalization of the matter power spectrum $\sigma_8$.
Those parameters are typically estimated by likelihood-free inference, which requires a function to predict the parameters from simulations.
% such as Approximate Bayesian computation (ABC)
As that is complicated to setup, prediction methods are typically benchmarked on the classification of spherical maps into cosmological model classes \citep{schmelze2017cosmologicalmodel}.
We used the same task, data, and setup as \citet{perraudin2019deepspherecosmo}: the classification of $720$ partial convergence maps made of $n \approx 10^6$ pixels ($1/12 \approx 8\%$ of a sphere at $N_{side} = 1024$) from two $\Lambda$CDM cosmological models, ($\Omega_m = 0.31$, $\sigma_8 = 0.82)$ and ($\Omega_m = 0.26$, $\sigma_8 = 0.91)$, at a relative noise level of $4$.
% 60 maps * 12 parts = 720 parts
Convergence maps represent the distribution of over- and under-densities of mass in the universe \citep[see][for a review of gravitational lensing]{bartelman2010gravitationallensing}.

@@

\begin{table}
    \centering
    \begin{tabular}{l c c c c}
		\toprule
		& accuracy & conv. equiv. error & inference \\ % & training time & NN inv. error
		\midrule
		2D CNN baseline \citep{perraudin2019deepspherecosmo} & \todo{??} &  \\ % &  & \todo{can we?}
		8 neighbors, CNN variant, \citep{perraudin2019deepspherecosmo} & \todo{??} & \todo{??} &  \\
		8 neighbors, FCN variant, \citep{perraudin2019deepspherecosmo} & 83.8 & \todo{??} & 187\,ms \\ % & 1\,h\,45\,m
		8 neighbors, optimal kernel width & 87.1 & \todo{??} & 185\,ms \\ % & 1\,h\,45\,m
		20 neighbors, optimal kernel width & 91.3 & \todo{??} & 250\,ms \\ % & 2\,h\,30\,m
		40 neighbors, optimal kernel width & 92.5 & \todo{??} & 363\,ms\\ % & 3\,h\,45\,m
		\bottomrule
    \end{tabular}
	\caption{
		Results on the classification of partial convergence maps.
		\todo{While extreme non-equivariance (the 2D CNN baseline) penalizes performance, improving equivariance doesn't improve performance.}
		\todo{add speed to show trade-off}
	}
    \label{tab:cosmo}
\end{table}

\todo{
}

Results are shown in \tabref{cosmo}.
\todo{
	While extreme non-equivariance (the 2D CNN baseline) penalizes performance, better equivariance doesn't improve performance.
	It might be that the lack of built-in equivariance is compensated by learned equivariance (as done when augmenting data during training).}
\todo{
	Alternative: DSv2 better than v1 -> equivariance has practical implications.
	Clear cost-accuracy tradeoff with the number of neighbors.
}
This experiment moreover demonstrates DeepSphere's flexibility (using partial spherical maps) and scalability (competing spherical CNNs were tested on maps of at most $10,000$ pixels).

>>>>

\usepackage{hyperref}
\usepackage{url}
\usepackage{breqn}
\usepackage{booktabs}
\usepackage{siunitx} % used in SHREC table
\usepackage{bbm}        % for mathbbm
\usepackage{graphicx} \graphicspath{{figures/}}
\usepackage[export]{adjustbox} % figure vertical alignment
%\usepackage{floatrow} % figure and table together
\usepackage{capt-of} % or \usepackage{caption} % figure and table together
\usepackage{subfig}
%\usepackage{wrapfig}
%\usepackage{floatflt}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{wrapfig}

@@

\todo{
\begin{itemize}
	\item shall we really name it? Or simply reference the general idea of a graph-based spherical CNN.
	\item adjacency matrix: $A$ instead of $W$?
	\item anonymize DeepSphere citation
% 	\item climate spherical figure: inputs with TC and AR labels superimposed [Frédérick]
	\item citet for inline, citep for parenthesis
	\item More details on related work? Not sure we have the space. If yes, detailed comparison with related work after method.
	\item PSD of climate data: how does it compare to cosmo and 3D objects?
	\item align the spheres in figure 1
	\item mention that \citep{esteves2018sphericalcnn} and \citep{cohen2018sphericalcnn} computed equivariance error
	\item check footnotes
	\item climate data on spectrum plot
	\item result tables: only what supports the main message in the text, everything we have in the appendix
	\item Unify vocabulary. Sampling, discretization, or pixelization of the sphere? Signal, maps, functions. Irregular, uneven. Coarsening, down-sampling.
	\item check English vs American English
	\item unify enumerations. I like (i), (ii), etc.
	\item notation: should we use indices from 1 to n instead of 0 to n-1 for compactness?
	\item add \citet{bruna2013gnn} in equivariance? They were the first to do a spherical CNN with graphs. Graph is built from a uniform random sampling of the sphere.
	\item figure placements
	\item do the experiments on the icosahedral grid with the pooling as defined in the paper
	\item put all the code (learning experiments from Frédérick and equivariance experiments from Martino) in a common reproducible git repository to be linked from the paper
	\item update pygsp and merge sphere branch
	\item check that NN architectures and hyper-params are correctly reported
	\item keep NN architecture and hyper-params in the main text? Move the paragraph to appendix if not enough space.
	%\item shall we still site adam and batch norm? no, mainstream enough
	\item check results for ModelNet40 and integrate in text
	\item Check that HEALPix is introduced in the introduction
	\item heuristic used to set the kernel width in v1. Add values to \figref{kernel_widths} for comparison.
	\item check experiments section [Frédérick]: all the results numbers and NN + training details
	\item compute the invariance error of the whole NNs (before and after training, invariance can be learned)
	\item add it with equivariance error of the convolution to \tabref{cosmo}
\end{itemize}
}
\newpage

@@

Given observations, cosmologists estimate the posterior probability of cosmological parameters, such as the matter density $\Omega_m$ and the normalization of the matter power spectrum $\sigma_8$.
Those parameters are typically estimated by likelihood-free inference, which requires a function to predict the parameters from simulations.
% such as Approximate Bayesian computation (ABC)
As that is complicated to setup, prediction methods are typically benchmarked on the classification of spherical maps into cosmological model classes \citep{schmelze2017cosmologicalmodel}.
We used the same task, data, and setup as \citet{perraudin2019deepspherecosmo}: the classification of $720$ partial convergence maps made of $n \approx 10^6$ pixels ($1/12 \approx 8\%$ of a sphere at $N_{side} = 1024$) from two $\Lambda$CDM cosmological models, ($\Omega_m = 0.31$, $\sigma_8 = 0.82)$ and ($\Omega_m = 0.26$, $\sigma_8 = 0.91)$, at a relative noise level of $3.5$ (i.e., the signal is hidden in noise of $3.5$ times higher power).
% 60 maps * 12 parts = 720 parts
Convergence maps represent the distribution of over- and under-densities of mass in the universe \citep[see][for a review of gravitational lensing]{bartelman2010gravitationallensing}.

@@

\begin{table}
	\begin{minipage}[b]{0.7\linewidth}
    \centering
    \begin{tabular}{l c c}
		\toprule
		& accuracy & time \\ % & training time & NN inv. error
		\midrule
		\citet{perraudin2019deepspherecosmo}, 2D CNN baseline & \todo{??} & \todo{??} \\
		\citet{perraudin2019deepspherecosmo}, CNN variant, $k=8$ & \todo{??} & \todo{??} \\
		\citet{perraudin2019deepspherecosmo}, FCN variant, $k=8$ & 83.8 & 187\,ms \\ % & 1\,h\,45
		$k=8$  neighbors, $t$ from \secref{optimal} & 87.1 & 185\,ms \\ % & 1\,h\,45
		$k=20$ neighbors, $t$ from \secref{optimal} & 91.3 & 250\,ms \\ % & 2\,h\,30
		$k=40$ neighbors, $t$ from \secref{optimal} & 92.5 & 363\,ms \\ % & 3\,h\,45
		\bottomrule
    \end{tabular}
	\caption{
		Results on the classification of partial convergence maps.
		Lower equivariance error translates to higher performance.
	} \label{tab:cosmo}
	\end{minipage} \hfill
	\begin{minipage}[b]{0.25\linewidth}
		\includegraphics[width=\linewidth]{cosmo_cost_accuracy_tradeoff}
		\captionof{figure}{Tradeoff between cost and accuracy.}
		\label{fig:cosmo:tradeoff}
	\end{minipage}
\end{table}

Results (\tabref{cosmo}) show that a lower equivariance error for convolution translates to higher performance.
There is a clear cost-accuracy tradeoff, controlled by the number of neighbors $k$ (\figref{cosmo:tradeoff}).
\todo{
	While extreme non-equivariance (the 2D CNN baseline) penalizes performance, better equivariance doesn't improve performance.
	It might be that the lack of built-in equivariance is compensated by learned equivariance (as done when augmenting data during training).}
\todo{
	Alternative: DSv2 better than v1 -> equivariance has practical implications.
}
This experiment moreover demonstrates DeepSphere's flexibility (using partial spherical maps) and scalability (competing spherical CNNs were tested on maps of at most $10,000$ pixels).
